Campaign Contributions From Ohio For Election 2012 
by Christian Villere
========================================================

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, 
                      message=FALSE)

# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for 
#this code chunk.# This prevents the code from 
#displaying in the knitted HTML output.You should set
#echo=FALSE for all code chunks in your file.
#install.packages("knitr", dependencies = T)
#install.packages("contrib.url")

library(RColorBrewer)
library(ggplot2)
library(dplyr)


```

```{r echo=FALSE, Load_the_Data}

setwd("C:/Users/Plo/exploratory Data Analysis/Final Project")

#creating file connection object
con <- file("P00000001-OH.csv")

#read in column names from file
ln <- readLines(file("P00000001-OH.csv"), n=1) 


#So, read column names separately, add one more dummy 
#column named "dummy"
ln <- paste (ln,'dummy', sep=',')

#strip column names from file, split the column names 
#into substrings, and save into memory 
col_names_raw <- unlist(strsplit(ln, ","))


#read csv skipping the first row and supplying 
#the col names
contribs_oh <- read.csv("P00000001-OH.csv", 
                        col.names=col_names_raw, 
                        header=FALSE, row.names=NULL,
                        skip=1)

#closing file connection
close(con)

#removing dummy variable, which was added in 
#order to read in the data. 
contribs_oh <- subset(contribs_oh, select = -dummy)


#removing negative or 0 entries, which are 
#bookkeeping items. 
contribs_ohio <- subset(contribs_oh, 
                        contribs_oh$contb_receipt_amt 
                        == 0)
contributions_olow <- subset(contribs_oh, 
                      contribs_oh$contb_receipt_amt < 0)

#This is the dataset I will work with 
#throughout the analysis. 
contribs_oh <- subset(contribs_oh, 
                      contribs_oh$contb_receipt_amt > 0)




str(contribs_oh)
                           
```

This report explores 151,479 campaign contributions 
made to Presidential Campaigns in the 2012 election 
from the state of Ohio. We will explore from where
in Ohio these contributions came from, any other supplied
demographic information that can help us shed light
on the contributors, which candidates had more successful 
fundraising efforts in the state, and the relationship
between those efforts and the outcome of the election.
Our target variable, or variable of interest, is 
contribution amount represented as 
contb_receipt_amt in the dataset.


# Univariate Plots Section
```{r echo=FALSE}

p0 <- qplot(x= contb_receipt_amt, data = 
              contribs_oh, binwidth = 20) 

p0 + scale_x_continuous(lim = c(0, 2000))
```

Data is heavily skewed toward lower contribution 
amounts. Lets take a log of the x scale to see if 
we can determine which of the contribution amounts 
was most popular. 


```{r echo=FALSE, Univariate_Plots}
p1 <- qplot(x= contb_receipt_amt, 
            data = contribs_oh, binwidth = .1) 

p1 + scale_x_log10(breaks = c(0, 10, 100, 1000))

```

The most common contribution amount is $100. That 
makes a lot of sense since its a common amount in 
general and on an online contribution form it is 
most likely a box that folks can check to determine 
that's how much they want to give. Due to this variable
looking so approximately normal, I'm going to use the 
scale transformation throughout the report. 


```{r echo=FALSE}
p2 <- qplot(x = cand_nm, data = contribs_oh) +
  theme(axis.text.x = element_text(angle = 45, 
                      hjust = 1, vjust = 1))

p2
```

Although there were 14 candidates in the race, 
only two, Mitt Romney and Barack Obama received 
a significant number of donations. The instructions 
suggest to add the gender of each candidate or the 
political party, but at this time neither additional 
variable appears interesting or enlightening.


```{r echo=FALSE}
contrib <- table(contribs_oh$contbr_city)
contrib_can <- sort(contrib)
top_cities <- tail(names(contrib_can), 9)
contrib_cities <- subset(contribs_oh, 
                         contbr_city %in% top_cities)

p3 <- qplot(x = contbr_city, data = contrib_cities) + 
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1, vjust = 1))

p3

head(summary(contribs_oh$contbr_city), 20)
```


Not terribly surprising. The top cities were 
the six I had heard of in Ohio, and four I had not.
Given that these donations seem to be coming from 
all over the state, and are not concentrated in just 
a couple places, it will be interesting to see if Obama
receives more or less total contributions in the state 
than the Republicans.



```{r echo=FALSE}
occupations <- table(contribs_oh$contbr_occupation)
occupations_oh <- sort(occupations)
top_occs <- tail(names(occupations_oh), 9)
top_occupations <- subset(contribs_oh, 
                      contbr_occupation %in% top_occs)

p4 <- qplot(x = contbr_occupation, 
            data = top_occupations) + 
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1, vjust = 1))

p4

head(summary(contribs_oh$contbr_occupation), 20)


```

Here's where things get interesting from my point 
of view. The occupation of the folks who gave most 
frequently is listed as retired. Let's delve 
into that a little bit. It is pretty amazing
they represent such a large number of the 
political contributions in the state. 
Judging from the summary of the professions, 
it appears this is a list of only individuals.If 
the list did include businesses, it would make the
representation of retired folks on this 
list even larger. Unfortunately, two of the 
professions listed, "information requested
per best effors" and "information requested" do
not provide us with any information about the 
profession of almost 10,000 contributors or 
dataset observations.


```{r echo=FALSE}
contribs_oh_retired <- subset(
  contribs_oh, contbr_occupation == "RETIRED")

p5 <- qplot(x = cand_nm, data = contribs_oh_retired) + 
  theme(axis.text.x = element_text(angle = 45, 
                                hjust = 1, vjust = 1))

p5

```

Unless there was a rash of young retirees in Ohio, 
its mildly surprising Obama had roughly 50% more 
contributions from that age group than Mitt Romney.
Older folks typically vote Republican. In a swing 
state like Ohio, President Obama was probably 
appealing to voters across the aisle with such
high numbers of contributions from this age group.
Let's take a look at where these retired contributors 
lived.


```{r echo=FALSE}
retired_oh <- table(contribs_oh_retired$contbr_city)
retired_oh <- sort(retired_oh)
top_retired <- tail(names(retired_oh), 12)
top_retired_city <- subset(contribs_oh_retired, 
                           contbr_city %in% top_retired)

p6 <- qplot(x = contbr_city, data = top_retired_city) + 
  theme(axis.text.x = element_text(angle = 45, 
                                  hjust = 1, vjust = 1))

p6


```


The top 3 cities for where political contributions 
came from is the same for retired people as it is 
for the entire sample of contributors. 



# Univariate Analysis

### What is the structure of your dataset?

The dataset consists of 18 variables of 151,479 
observations or contributions made to 
Presidential campaigns in Ohio for the 2012 
Presidential election. Almost all the variables are 
factors, with the exception of contb_receipt_dt, 
contb_receipt_amt, and file_num. 

### What is/are the main feature(s) of interest in your dataset?
The main features of interest in the dataset are the 
contribution amount, and the candidate awarded 
with that contribution.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
Other features of interest are what the contributors 
do for a living and where they actually live.  

### Did you create any new variables from existing variables in the dataset?
I didn't create any new variables because I can't 
determine there is any benefit to doing so at this 
stage. The assignment suggests adding a variable for 
candidate gender or candiate political party. 
Right now, I don't see any analytical benefit 
to adding either of those two variables or any 
other variable. 

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?
There were a couple things I had to do early on 
that were unique to this dataset. First, it was not 
assimple as using the read.csv command to read in the 
data. According to another user, whom I credit in 
my resources section, the 2012 data has a trailing comma. 
Due to this, R thinks there is an extra column. I used 
some code from the user to create the additional 
column, read in the data, and manage to do so 
without causing any issues with the column names 
being moved one to the right, which was a problem 
with other solutions. Second, it appears there 
were some contribution amounts of zero. I subsetted 
them from the data, due to them not being necessary,
and may look at them later on depending where the 
analysis takes me.  


# Bivariate Plots Section

Let's take a look at contribution amount and 
contributor city together.

```{r echo=FALSE}
p7 <- ggplot(aes(x = contbr_city, y = contb_receipt_amt),
      data = contrib_cities) + 
      theme(axis.text.x = 
      element_text(angle = 45, hjust = 1, vjust = 1)) + 
      geom_boxplot() 


p7 + scale_y_log10()

```

Interesting that the more populous urban cities have more
donations, Dublin has the highest median. The 
interquartile ranges are relatively similar. Dayton and 
and Columbus have the most numerous amount of outliers 
in the graph. 

Let's consider contribution amount and occupation


```{r echo=FALSE}
p8 <- ggplot(aes(x = contbr_occupation,
  y = contb_receipt_amt), data = top_occupations) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, 
  vjust = 1)) + geom_boxplot()


p8 + scale_y_log10()

```

interestingly, those without an attributed occupation
have the highest median contribution amount. Retired 
folks have many more outliers, but of smaller amounts,
while attorney, homemaker, and 
physican have less outliers, but at larger amounts.
The range of amounts for teachers is quite small, 
but several of them gave atleast $500. That's pretty
impressive since its considered a poorly 
compensated occupation.


Let's consider contribution amount and being retired.


```{r echo=FALSE}
p9 <- ggplot(aes(x = contbr_occupation, 
  y = contb_receipt_amt), data = contribs_oh_retired) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1,
  vjust = 1)) + geom_boxplot() 

p9 + scale_y_log10(breaks = c(0, 10, 100, 1000))

summary(contribs_oh_retired$contb_receipt_amt)
```

Fairly plain graph. Median line sits right at $50.
Top of the box at $100, and lower end of the box at 
$25. Lots of outliers beyond the $1000 mark. 


Let's take a look at contribution amount and date


```{r echo=FALSE}
#converting date from factor to a date format
contribs_oh$contb_receipt_dt <- 
  as.Date(contribs_oh$contb_receipt_dt, "%d-%b-%y")

#group by month
contribs_oh$month <- 
  as.Date(cut(contribs_oh$contb_receipt_dt,
  breaks = "month"))

#group by week
contribs_oh$week <- 
  as.Date(cut(contribs_oh$contb_receipt_dt,
  breaks = "week",
  start.on.monday = FALSE))


p11 <- ggplot(aes(x = contb_receipt_dt, 
  y = contb_receipt_amt), data = contribs_oh) + 
  theme(axis.text.x = element_text(angle = 45, 
  hjust = 1, vjust = 1)) + geom_jitter(alpha = 1 / 35)

p11 + scale_y_log10()

```

Not much of a trend in contribution amount and date. 
The amounts appear to stay constant across time. 
There does appear to be more data points 
toward the end of the election season. I want to try
grouping the dates together to see if I can spot
any kind of a trend in the amounts 
given in each month. 


```{r echo=FALSE}

month_contributions <- group_by(contribs_oh, month)
monthly.mon_total <- summarise(month_contributions, 
             monthly_total = sum(contb_receipt_amt),
             monthly_mean = mean(contb_receipt_amt),
             n = n())

monthly.mon_total <- arrange(monthly.mon_total)

p12 <- ggplot(aes(x = month, y = contb_receipt_amt),
  data = contribs_oh) + theme(axis.text.x = 
  element_text(angle = 45, hjust = 1, vjust = 1)) + 
  geom_boxplot(aes(group = month)) + 
  ylim(0, 2500)

p12

p12 + scale_y_log10()
```

I want to show both the log transformed graph and not 
transformed graph here. In the former, you can
begin to see a trend in when the contributions come 
to a campaign. It looks like the contributions ramp 
up a little bit after the first month of 2012, and 
around the conventions. In the latter, 
you can see a more broad trend of a higher number
of outliers after the first of the year.


```{r echo=FALSE}
p13 <- ggplot(aes(x = month, y = monthly_total), 
  data = monthly.mon_total) + theme(axis.text.x = 
  element_text(angle = 45, hjust = 1, vjust = 1)) + 
  geom_line()

p13 


p14 <- ggplot(aes(x = month, y = monthly_total), 
  data = monthly.mon_total) + theme(axis.text.x = 
  element_text(angle = 45, hjust = 1, vjust = 1)) + 
  geom_bar(stat="identity")

p14

```

Contributions definitely ramp up toward the end
of the election per the monthly totals. That 
makes a good bit of sense as there is a huge 
push to get donations toward the latter part 
of a campaign. Also voter interest tends to see 
a surge in the last few months of a campaign. 


# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?
One interesting relationship I noticed was the somewhat
generous giving of teachers in the state of Ohio. The 
interquartile range is on the smaller end as it 
compares to the other top occupations; 
however there are still several outliers on the 
lower end of the y-axis range. This pattern 
actually holds true with all of the occupations: 
smallish interquartile ranges with many outliers. 
In regard to contribution amount and city, there 
were more contributors and more outliers in the higher 
populated urban cities.



### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

I did not observe any interesting relationships 
between other features of the dataset. However, 
I did create a couple new variables to look at 
the relationship of contribution amount totals 
across time, month and week.


### What was the strongest relationship you found?

The strongest relationship I was able to find 
was between the total amount given in a month,
and the month of the campaign. The monthly totals 
given seemed to ramp up as election day got closer. 



# Multivariate Plots Section

```{r echo=FALSE}
  

cand_contributions <- group_by(contribs_oh, 
                               cand_nm, month)
cand_contributions <- 
  subset(cand_contributions, cand_nm %in% 
  c("Obama, Barack", "Romney, Mitt", 
    "Santorum, Rick", "Paul, Ron"))

cand.mon_total <- summarise(cand_contributions, 
             monthly_total = sum(contb_receipt_amt),
             monthly_mean = mean(contb_receipt_amt),
             n = n())

summary(cand.mon_total)


p15 <- ggplot(aes(x = month, y = monthly_total), 
  data = cand.mon_total) + theme(axis.text.x = 
  element_text(angle = 45, hjust = 1, vjust = 1)) + 
  geom_line(aes(color = cand_nm), size = 1.20) +
  scale_color_brewer(type = "qual", palette = "Accent",
  guide = guide_legend(title = "Candidate", 
  reverse = T)) + theme_dark()

p15


p15a <- ggplot(aes(x = month, y = contb_receipt_amt),
  data = cand_contributions) + geom_line(aes(color = cand_nm),
  stat = 'summary', fun.y = mean, size = 1.15) +
  geom_line(linetype = 3, stat = "summary", fun.y = mean) +
  scale_color_brewer(type = "qual", palette = 'Dark2', 
  guide = guide_legend(title = "Candidate", reverse = T)) +
  theme_light()

p15a

```

Monthly totals for Romney are higher in almost 
every month. The grand mean follows a mostly similar 
trend to the rest of the averages for the top 4 candidates. 


```{r echo=FALSE}  
city_contributions <- group_by(contribs_oh, 
                               contbr_city, month)

city_contributions <- subset(city_contributions,
                             contbr_city %in% top_cities)

city.mon_total <- summarise(city_contributions, 
             monthly_total = sum(contb_receipt_amt),
             monthly_mean = mean(contb_receipt_amt),
             n = n())

summary(city.mon_total)

p16 <- ggplot(aes(x = month, y = monthly_total),
  data = city.mon_total) + 
  theme(axis.text.x = 
  element_text(angle = 45, hjust = 1, vjust = 1)) + 
  geom_line(aes(color = contbr_city), size = 1.20) +
  scale_color_brewer(type = "qual", palette = "Set1", 
  guide = guide_legend(title = "City", reverse = T)) +
  theme_dark()

p16

p16a <- ggplot(aes(x = month, y = contb_receipt_amt),
  data = city_contributions) +
  geom_line(aes(color = contbr_city), stat = 'summary',
  fun.y = mean, size = 1.15) +
  geom_line(linetype = 3, stat = "summary", 
  fun.y = mean) +
  scale_color_brewer(type = "qual", palette = "Set3",
  guide = guide_legend(title = "City", reverse = T)) +
  theme_dark()

p16a + scale_y_log10()

```

Monthly totals for contributions rack up in a fairly
predictable fashion. The larger urban areas have
higher monthly totals. In terms of averages, the plot 
lines are a bit more scattered, but for the 
most part follow the same trend of starting high, moving
lower, before spiking in January 2012, moving lower, 
then spiking again around the conventions, and then 
leveling off for good.


```{r echo=FALSE}
occs_contributions <- group_by(contribs_oh, 
                    contbr_occupation, month)
occs_contributions <- subset(occs_contributions, 
                    contbr_occupation %in% top_occs)

occs.mon_total <- summarise(occs_contributions, 
             monthly_total = sum(contb_receipt_amt),
             monthly_mean = mean(contb_receipt_amt),
             n = n())

summary(occs.mon_total)

p17 <- ggplot(aes(x = month, y = monthly_total), 
  data = occs.mon_total) + theme(axis.text.x = 
  element_text(angle = 45, hjust = 1, vjust = 1)) + 
  geom_line(aes(color = contbr_occupation), 
  size = 1.20) + scale_color_brewer(type = "qual", 
  palette = "Set1",guide = guide_legend(
  title = "Occupaton", reverse = T)) + theme_dark()

p17



p17b <- ggplot(aes(x = month, y = contb_receipt_amt),
  data = occs_contributions) + geom_line(aes(color = 
  contbr_occupation), stat = 'summary', 
  fun.y = mean, size = 1.15) + geom_line(linetype = 3,
  stat = "summary", fun.y = mean) + 
  scale_color_brewer(type = "qual", palette = "Set1",
  guide = guide_legend(title = "Occupation", 
  reverse = T)) +
  theme_dark()

p17b + scale_y_log10()


```

It shouldn't be a huge surprise that retired folks
have a much higher monthly total in the later
months of a campaign when engagement is the highest.
They are the most represented group in the 
dataset so it makes pretty good sense that 
they would consisently have higher monthly 
contribution totals compared to other "professions." 

Before I wrap up the multivariate plots section, I want 
to see how the top occupations and the top cities
gave to the top candidates. To do this, I'll create
a couple new dataframes using dplyr.


```{r echo=FALSE}
#let's look at cities first

contrib_cities_cand <- 
  subset(contrib_cities, cand_nm %in% 
  c("Obama, Barack", "Romney, Mitt", 
    "Santorum, Rick", "Paul, Ron"))

contrib_cand_select <- select(contrib_cities_cand,
                              contbr_city, cand_nm,
                              contb_receipt_amt)

cities_cand_totals <- group_by(contrib_cand_select,
                               contbr_city, cand_nm) %>%
  summarise(mean = mean(contb_receipt_amt), 
            sum = sum(contb_receipt_amt),
            n = n())

P25 <- ggplot(cities_cand_totals, aes(cand_nm, 
      contbr_city, fill = sum)) + geom_raster() +
  scale_fill_gradientn(colours = c("#0000FFFF","#FFFFFFFF","#FF0000FF"))

P25

```

Rick Santorum did not receive any contributions in Shaker Heights. 
Mitt Romney raked in the dough in Cinncinati. Barack Obama got his
contributions from a spread of cities.

Now lets see what the graph looks like for occupations. 


```{r echo=FALSE}

occs_contributions <- subset(occs_contributions, 
                    contbr_occupation %in% top_occs)


contrib_occs_cand <- 
  subset(occs_contributions, cand_nm %in% 
  c("Obama, Barack", "Romney, Mitt", 
    "Santorum, Rick", "Paul, Ron"))


occup_cand_select <- select(contrib_occs_cand,
                     contbr_occupation, cand_nm,
                              contb_receipt_amt)
occup_cand_select <- subset(occup_cand_select, 
                            select = -month)
                             

occup_cand_totals <- group_by(occup_cand_select,
                               contbr_occupation, 
                               cand_nm) %>%
  summarise(mean = mean(contb_receipt_amt), 
            sum = sum(contb_receipt_amt),
            n = n())

P25b <- ggplot(occup_cand_totals, 
  aes(contbr_occupation, cand_nm, fill = sum)) +
  geom_raster() + 
  scale_fill_gradientn(colours = 
  c("#DADA1B","#2EDA1B","#B0183B")) +
   theme(axis.text.x = element_text(angle = 45, 
                      hjust = 1, vjust = 1))

P25b
```

Mitt Romney cleaned up with retired folks and those 
for whom information on occupation could not be 
obtained. He also did well with homemakers. 
Barack Obama did well with retired folks as well, 
and had support across almost all the 
occupations.




# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

In looking at averages of contribution amounts by 
month for the candidates, the averages were higher
earlier on in the campaign compared to later. This
makes sense because those really engaged politically 
tend to be of a higher socio-economic status, who 
give their money to their preferred 
campaign earlier on, while the other end of the 
spectrum in terms of class comes around later on,
and is of a lower socio-economic status. 

Across month, city, and occupation, the monthly 
contribution totals increased as the campaign drew 
near. This makes sense because more people are 
tuning in. These totals are going up, despite the
fact that average campaign contributions seem to 
decrease as the election gets closer. That's an 
interesting, yet expected dicotmy. 


### Were there any interesting or surprising interactions between features?
The most surprising or interesting interaction
I saw was between the month, monthly_total, and 
candidate. Despite higher contribution totals 
for Romney in the state, he was soundly defeated
by Obama in the election.


------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}


p1a <- qplot(x= contb_receipt_amt, data = contribs_oh,
  binwidth = .1) +
  scale_x_log10(breaks = c(0, 10, 100, 1000)) + 
  labs(title = "Log10 Individual Contribution Amount",
  x = "Contribution Amount (in USD)",
  y = "Number of Contributions")

p1a

  

```

### Description One

The distribution of individual contributions on a
log scale is fairly normalized. There looks to be
slightly higher frequency on the lower side of $100,
than on the higher side. Also, amounts 10, 50, 100,
500, and 1000, are pretty well represented. 


### Plot Two
```{r echo=FALSE, Plot_Two}

p8b <- ggplot(aes(x = contbr_occupation, 
  y = contb_receipt_amt), data = top_occupations) + 
  theme(axis.text.x = element_text(angle = 45,
  hjust = 1, vjust = 1)) + geom_boxplot(fill = "red") +
  theme(panel.background = 
  element_rect(fill = 'green')) + 
  labs(title = "Contributions of Most Common Occupations",
  x = "Occupation", y = "Receipt Amount (in USD)") +
  scale_y_log10()

p8b

```

### Description Two

The interquartile range for most of the occupations
is fairly small. The interquartile range is 
slightly larger for occupations with a history of 
being highly compensated like Attorney or Physican.
I guess it should be taken for granted, but the number
of outlier contributions by those who are retired 
slightly surprised me. They are almost 1/5 of the 
data observations, and a good deal of them are 
giving more than the most common contribution of $100. 


### Plot Three
```{r echo=FALSE, Plot_Three}

p15b <- ggplot(aes(x = month, y = contb_receipt_amt),
  data = cand_contributions) +
  geom_line(aes(color = cand_nm), stat = 'summary', 
  fun.y = mean, size = 1.15) +
  geom_line(linetype = 3, stat = "summary", 
  fun.y = mean) +
  scale_color_brewer(type = "qual", 
  guide = guide_legend(title = "Candidate", 
  reverse = T)) +
  labs(title = "Average Contribution By Month and 
  Candidate", x = "Month of The Year", 
  y = "Average Contribution (in USD)") +
  theme(panel.background = 
  element_rect(fill = 'green'))

p15b


```

### Description Three

It's somewhat surprising that the trend for the
average contribution goes down as the election 
gets closer. From left to right, the averages 
begin to move downward, there is a spike at the
beginning of the year, before they begin to 
decrease again after the time around the 
conventions. It sort of makes sense given those
most likely to be engaged early on are wealthier
and able to give more. Those who are less engaged
tend to be less wealthy and more likely to engage
with the race toward the latter weeks of a campaign,
and there many more of those folks compared to the
earlier crowd. 

------

# Reflection

It was a pleasure to be able to work on this dataset.
As I reflect on the project as a whole, the main 
difficulty stemmed from wondering how I was going to
be able to present any findings of use or insights, 
which may not seem plausible on the surface. The data
is somewhat limited because a good many of the 
variables are administrative in nature or for 
bookkeeping reasons, such as form_id. However, I was
able to add a time series element to the data by adding
new variables, month and Week, to the dataset. This 
led to some counterintuitive conclusions I myself 
wasn't expecting. For example, in regard to Mitt
Romney, fewer people in a higher concentration 
of cities, gave more fundraising dollars to him 
than Obama, who received his contributions from 
more people spread in concentration across
more cities.

As for ideas or questions for future work, I think it
would be worthwhile to do a deeper dive into the 
contributors/contributions for Mitt Romney and the
contributors/contributions for Barack Obama. Even 
though Romney received more money and higher money
totals, he was beaten in the state by Obama. This 
suggests that campaigns aren't won by the amount of
money a candidate receives, but rather the number of
people who give to the candidate. Did Barack Obama 
have more contributors than Mitt Romney in the state
of Ohio? If so, what were the demographics of those 
people? What does it suggest about the difference 
between the two campaigns?

I can't say there were any really important decisions
regarding the analysis of the dataset. However, early 
on, I did key in on the variables in the dataset I 
thought could illuminate the variable of interest, 
contribution amount, the best without doing a bunch
of investigation on many of the variables I did not
use, because they did not seem like they would add
anything to the analysis. Also, I keyed in on the 
top occupations, cities, and candidates without doing
much analysis of the middle or the bottom in each of
those categories. The decision to do that was mainly
due to thinking it would be harder to find any analysis
or trend in a graph of any of those variables or 
attributes of the dataset. 









